### 2019-12-11 Statistical Methods/03_graphical-summaries.rmd
statistical methods, ch 3
statistical methods, ch 3

For the data in Table 2.1, dealing with changes in cholesterol levels, create a histogram with the R function hist.

```{r, echo = T}
data <- read.csv(paste0(data.dir, "ibtable2_1_dat.txt"), header = F)

@@

For the data in Table 2.2, create a histogram using the R function _hist_ and speculate about whether values less than zero are outliers.

```{r, echo = T}
data <- read.csv(paste0(data.dir, "ibtable2_2_dat.txt"), header = F)

@@

@@

@@

```{r, echo = T}
x <- c(121, 124, 129, 129, 130, 130, 131, 131, 132, 132, 132, 133, 133, 134, 134, 134,
134, 135, 136, 136, 136, 136, 136, 137, 137, 138, 138, 138, 140, 143)

@@

@@

@@

For the dta in the previous exercise, verify that the classic outlier detection rule, given by _2.6_, finds three outliers, which match the values flagged as outliers by the MAD-median rule.

@@

>>>>

For the data in Table 2.1, dealing with changes in cholesterol levels, create a histogram with R.

```{r, echo = T, , fig.width=8, fig.height=4}
data <- data.table::fread(paste0(data.dir, "ibtable2_1_dat.txt"), fill = T)

suppressWarnings({print(
   ggplot(data.table(value = as.vector(as.matrix(data))), aes(value)) +
      geom_histogram(aes(fill = ..count..), bins = 30)
)})

@@

For the data in Table 2.2, create a histogram using R, and speculate about whether values less than zero are outliers.

```{r, echo = T, fig.width=8, fig.height=4}
data <- data.table::fread(paste0(data.dir, "ibtable2_2_dat.txt"), fill = T)

values <- data.table(value = as.vector(as.matrix(data)))
values <- values[!is.na(values$value)]

suppressWarnings({print(
   ggplot(values, aes(value)) +
      geom_histogram(aes(fill = ..count..), bins = 30)
)})

@@

Below -45 & above 50 appear to be outliers.

@@

```{r, echo = T, fig.width=8, fig.height=3.5}
# MAD outliers

values[ abs(values$value - median(values$value)) / mad(values$value) > 2.27 ]

# Classical Outliers

values[ abs(values$value - mean(values$value)) / sd(values$value) > 2]
```

@@

```{r, echo = T, fig.width=8, fig.height=4}
x <- c(121, 124, 129, 129, 130, 130, 131, 131, 132, 132, 132, 133, 133, 134, 134, 134,
134, 135, 136, 136, 136, 136, 136, 137, 137, 138, 138, 138, 140, 143)

@@

ggplot(data.table(value = x), aes(x)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_density(fill = "darkorange", alpha = .35)

@@

x[ abs(x - median(x)) / mad(x) > 2.27 ]

# Classical Outliers

x[ abs(x - mean(x)) / sd(x) > 2]

@@

For the data in the previous exercise, verify that the classic outlier detection rule, given by _2.6_, finds three outliers, which match the values flagged as outliers by the MAD-median rule.

_Confirmed 3 outliers with both rules_

@@

_There appears to be masking going on, as there are clearly more than 3 outliers._
